---
title: "Multivariate, HomeWork II"
author: "Habib ezatabadi"
date: '2022-10-06'
output:
  pdf_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = F, warning = F)
```



## loading libraries

```{r}
library(tidyverse)
library(MVA)
library(HSAUR3)
library(corrplot)
library(car)
dat <- USairpollution
```



## get a head of data and name of variables


```{r}

head(dat)
dim(dat)

glimpse(dat)
```


***
***


## Ex 2.1: Use the bivariate boxplot on the scatterplot of each pair of variables in the air pollution data to identify any outliers. Calculate the correlation between each pair of variables using all the data and the data with any identified outliers removed. Comment on the results.



```{r fig.width = 7, fig.height = 7}

cor_full <- cor(dat) %>% round(3)
cor_editted <- matrix(1, 7, 7, dimnames = list(names(dat), names(dat)))
Id <- 1:nrow(dat)
labs <- names(dat)

### 1

bvbox(cbind(dat[, 1], dat[, 2]), 
xlab = labs[1], ylab = labs[2], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 1], y = dat[, 2], labels = Id, col = "red")
ind1 <- c(7, 23, 30, 33)
cor_editted[1, 2] <- cor_editted[2, 1] <- cor(dat[-ind1, c(1, 2)])[1, 2] %>%
  round(3)

### 2

bvbox(cbind(dat[, 1], dat[, 3]), 
xlab = labs[1], ylab = labs[3], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 1], y = dat[, 3], labels = Id, col = "red")
ind2 <- c(14, 9, 38, 30, 33, 36)
cor_editted[1, 3] <- cor_editted[3, 1] <- cor(dat[-ind2, c(1, 3)])[1, 2] %>%
  round(3)




### 3

bvbox(cbind(dat[, 1], dat[, 4]), 
xlab = labs[1], ylab = labs[4], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 1], y = dat[, 4], labels = Id, col = "red")
ind3 <- c(7, 9, 16, 14, 30, 33)
cor_editted[1, 4] <- cor_editted[4, 1] <- cor(dat[-ind3, c(1, 4)])[1, 2] %>%
  round(3)



### 4

bvbox(cbind(dat[, 1], dat[, 5]), 
xlab = labs[1], ylab = labs[5], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 1], y = dat[, 5], labels = Id, col = "red")
ind4 <- c(7, 33, 40)
cor_editted[1, 5] <- cor_editted[5, 1] <- cor(dat[-ind4, c(1, 5)])[1, 2] %>%
  round(3)




### 5

bvbox(cbind(dat[, 1], dat[, 6]), 
xlab = labs[1], ylab = labs[6], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 1], y = dat[, 6], labels = Id, col = "red")
ind5 <- c(7, 31, 33, 2, 23)
cor_editted[1, 6] <- cor_editted[6, 1] <- cor(dat[-ind5, c(1, 6)])[1, 2] %>%
  round(3)




### 6


bvbox(cbind(dat[, 1], dat[, 7]), 
xlab = labs[1], ylab = labs[7], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 1], y = dat[, 7], labels = Id, col = "red")
ind6 <- c(2, 5, 33, 7, 31)
cor_editted[1, 7] <- cor_editted[7, 1] <- cor(dat[-ind6, c(1, 7)])[1, 2] %>%
  round(3)




### 7


bvbox(cbind(dat[, 2], dat[, 3]), 
xlab = labs[2], ylab = labs[3], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 2], y = dat[, 3], labels = Id, col = "red")
ind7 <- c(30, 7, 14)
cor_editted[2, 3] <- cor_editted[3, 2] <- cor(dat[-ind7, c(2, 3)])[1, 2] %>%
  round(3)




### 8


bvbox(cbind(dat[, 2], dat[, 4]), 
xlab = labs[2], ylab = labs[4], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 2], y = dat[, 4], labels = Id, col = "red")
ind8 <- c(30, 7, 14, 16)
cor_editted[2, 4] <- cor_editted[4, 2] <- cor(dat[-ind8, c(2, 4)])[1, 2] %>%
  round(3)



### 9


bvbox(cbind(dat[, 2], dat[, 5]), 
xlab = labs[2], ylab = labs[5], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 2], y = dat[, 5], labels = Id, col = "red")
ind9 <- c(31, 23, 16)
cor_editted[2, 5] <- cor_editted[5, 2] <- cor(dat[-ind9, c(2, 5)])[1, 2] %>%
  round(3)



### 10


bvbox(cbind(dat[, 2], dat[, 6]), 
xlab = labs[2], ylab = labs[6], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 2], y = dat[, 6], labels = Id, col = "red")
ind10 <- c(2, 12, 31, 23, 27)
cor_editted[2, 6] <- cor_editted[6, 2] <- cor(dat[-ind9, c(2, 6)])[1, 2] %>%
  round(3)




### 11


bvbox(cbind(dat[, 2], dat[, 7]), 
xlab = labs[2], ylab = labs[7], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 2], y = dat[, 7], labels = Id, col = "red")
ind10 <- c(2, 31, 23)
cor_editted[2, 7] <- cor_editted[7, 2] <- cor(dat[-ind10, c(2, 7)])[1, 2] %>% round(3)






### 12


bvbox(cbind(dat[, 3], dat[, 4]), 
xlab = labs[3], ylab = labs[4], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 3], y = dat[, 4], labels = Id, col = "red")
ind12 <- c(9, 14, 30, 16, 7)
cor_editted[3, 4] <- cor_editted[4, 3] <- cor(dat[-ind12, c(3, 4)])[1, 2] %>% round(3)





### 13


bvbox(cbind(dat[, 3], dat[, 5]), 
xlab = labs[3], ylab = labs[5], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 3], y = dat[, 5], labels = Id, col = "red")
ind13 <- c(30, 7, 40)
cor_editted[3, 5] <- cor_editted[5, 3] <- cor(dat[-ind13, c(3, 5)])[1, 2] %>% round(3)






### 14


bvbox(cbind(dat[, 3], dat[, 6]), 
xlab = labs[3], ylab = labs[6], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 3], y = dat[, 6], labels = Id, col = "red")
ind14 <- c(30, 7, 2)
cor_editted[3, 6] <- cor_editted[6, 3] <- cor(dat[-ind14, c(3, 6)])[1, 2] %>% round(3)





### 15


bvbox(cbind(dat[, 3], dat[, 7]), 
xlab = labs[3], ylab = labs[7], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 3], y = dat[, 7], labels = Id, col = "red")
ind15 <- c(30, 7, 31, 9)
cor_editted[3, 7] <- cor_editted[7, 3] <- cor(dat[-ind15, c(3, 7)])[1, 2] %>% round(3)




### 16


bvbox(cbind(dat[, 4], dat[, 5]), 
xlab = labs[4], ylab = labs[5], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 4], y = dat[, 5], labels = Id, col = "red")
ind16 <- c(30, 7, 14, 40)
cor_editted[4, 5] <- cor_editted[5, 4] <- cor(dat[-ind16, c(4, 5)])[1, 2] %>% round(3)







### 17


bvbox(cbind(dat[, 4], dat[, 6]), 
xlab = labs[4], ylab = labs[6], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 4], y = dat[, 6], labels = Id, col = "red")
ind17 <- c(30, 7, 14, 2)
cor_editted[4, 6] <- cor_editted[6, 4] <- cor(dat[-ind17, c(4, 6)])[1, 2] %>% round(3)



### 18


bvbox(cbind(dat[, 4], dat[, 7]), 
xlab = labs[4], ylab = labs[7], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 4], y = dat[, 7], labels = Id, col = "red")
ind18 <- c(30, 7, 14, 31)
cor_editted[4, 7] <- cor_editted[7, 4] <- cor(dat[-ind18, c(4, 7)])[1, 2] %>% round(3)




### 19


bvbox(cbind(dat[, 5], dat[, 6]), 
xlab = labs[5], ylab = labs[6], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 5], y = dat[, 6], labels = Id, col = "red")
ind19 <- c(2, 12, 31)
cor_editted[5, 6] <- cor_editted[6, 5] <- cor(dat[-ind19, c(5, 6)])[1, 2] %>% round(3)



### 20


bvbox(cbind(dat[, 5], dat[, 7]), 
xlab = labs[5], ylab = labs[7], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 5], y = dat[, 7], labels = Id, col = "red")
ind20 <- c(5, 31, 40)
cor_editted[5, 7] <- cor_editted[7, 5] <- cor(dat[-ind20, c(5, 7)])[1, 2] %>% round(3)




### 21


bvbox(cbind(dat[, 6], dat[, 7]), 
xlab = labs[6], ylab = labs[7], col = adjustcolor("blue", .5), pch = 16)
text(x = dat[, 6], y = dat[, 7], labels = Id, col = "red")
ind21 <- c(2, 12, 31)
cor_editted[6, 7] <- cor_editted[7, 6] <- cor(dat[-ind21, c(6, 7)])[1, 2] %>% round(3)


corrplot(cor_full, method = "number", 
col = c("orange", "green", "skyblue"), 
title = "correlation matrix with full data", type = "full", 
mar = c(0, 0, 5, 0), 
bg = "gray40", addgrid.col = "yellow")

corrplot(cor_editted, method = "number", 
col = c("orange", "green", "skyblue"), 
title = "correlation matrix with truncated data", type = "full", 
mar = c(0, 0, 5, 0), bg = "gray40", addgrid.col = "yellow")
```




***
***





## Ex. 2.2 Compare the chi-plots with the corresponding scatterplots for each pair of variables in the air pollution data. Do you think that there is any advantage in the former?

```{r fig.height = 7, fig.width = 7}


# 1

chiplot(x = dat[, 1], y = dat[, 2], pch = 16, col = "darkblue")

scatterplot(x = dat[, 1], y = dat[, 2], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[1], ylab = labs[2])




# 2


chiplot(x = dat[, 1], y = dat[, 3], pch = 16, col = "darkblue")

scatterplot(x = dat[, 1], y = dat[, 3], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[1], ylab = labs[3])









# 3


chiplot(x = dat[, 1], y = dat[, 4], pch = 16, col = "darkblue")

scatterplot(x = dat[, 1], y = dat[, 4], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[1], ylab = labs[4])




# 4


chiplot(x = dat[, 1], y = dat[, 5], pch = 16, col = "darkblue")

scatterplot(x = dat[, 1], y = dat[, 5], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[1], ylab = labs[5])




# 5


chiplot(x = dat[, 1], y = dat[, 6], pch = 16, col = "darkblue")

scatterplot(x = dat[, 1], y = dat[, 6], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[1], ylab = labs[6])




# 6


chiplot(x = dat[, 1], y = dat[, 7], pch = 16, col = "darkblue")

scatterplot(x = dat[, 1], y = dat[, 7], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[1], ylab = labs[7])




# 7


chiplot(x = dat[, 2], y = dat[, 3], pch = 16, col = "darkblue")

scatterplot(x = dat[, 2], y = dat[, 3], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[2], ylab = labs[3])




# 8


chiplot(x = dat[, 2], y = dat[, 4], pch = 16, col = "darkblue")

scatterplot(x = dat[, 2], y = dat[, 4], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[2], ylab = labs[4])




# 9


chiplot(x = dat[, 2], y = dat[, 5], pch = 16, col = "darkblue")

scatterplot(x = dat[, 2], y = dat[, 5], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[2], ylab = labs[5])




# 10


chiplot(x = dat[, 2], y = dat[, 6], pch = 16, col = "darkblue")

scatterplot(x = dat[, 2], y = dat[, 6], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[2], ylab = labs[6])




# 11


chiplot(x = dat[, 2], y = dat[, 7], pch = 16, col = "darkblue")

scatterplot(x = dat[, 2], y = dat[, 7], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[2], ylab = labs[7])




# 12


chiplot(x = dat[, 3], y = dat[, 4], pch = 16, col = "darkblue")

scatterplot(x = dat[, 3], y = dat[, 4], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[3], ylab = labs[4])




# 13


chiplot(x = dat[, 3], y = dat[, 5], pch = 16, col = "darkblue")

scatterplot(x = dat[, 3], y = dat[, 5], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[3], ylab = labs[5])




# 14


chiplot(x = dat[, 3], y = dat[, 6], pch = 16, col = "darkblue")

scatterplot(x = dat[, 3], y = dat[, 6], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[3], ylab = labs[6])




# 15


chiplot(x = dat[, 3], y = dat[, 7], pch = 16, col = "darkblue")

scatterplot(x = dat[, 3], y = dat[, 7], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[3], ylab = labs[7])




# 16


chiplot(x = dat[, 4], y = dat[, 5], pch = 16, col = "darkblue")

scatterplot(x = dat[, 4], y = dat[, 5], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[4], ylab = labs[5])




# 17


chiplot(x = dat[, 4], y = dat[, 6], pch = 16, col = "darkblue")

scatterplot(x = dat[, 4], y =  dat[, 6], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[4], ylab = labs[6])




# 18


chiplot(x = dat[, 4], y = dat[, 7], pch = 16, col = "darkblue")

scatterplot(x = dat[, 4], y = dat[, 7], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[4], ylab = labs[7])




# 19


chiplot(x = dat[, 5], y = dat[, 6], pch = 16, col = "darkblue")

scatterplot(x = dat[, 5],y = dat[, 6], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[5], ylab = labs[6])




# 20


chiplot(x = dat[, 5], y = dat[, 7], pch = 16, col = "darkblue")

scatterplot(x = dat[, 5], y = dat[, 7], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[5], ylab = labs[7])




# 21


chiplot(x = dat[, 6], y = dat[, 7], pch = 16, col = "darkblue")

scatterplot(x = dat[, 6], y = dat[, 7], pch = 16, col = "darkblue", 
            cex = 1.5, regLine = list(method = lm, lwd = 2, 
            col = adjustcolor("red", .5)), smooth = FALSE, 
            xlab = labs[6], ylab = labs[7])
```



***
***





## Ex. 2.3 Construct a scatterplot matrix of the body measurements data that
has the appropriate boxplot on the diagonal panels and bivariate boxplots
on the other panels. Compare the plot with Figure 2.17, and say which
diagram you find more informative about the data.


```{r fig.height = 9, fig.width = 9}

library(httpgd)
hgd()
hgd_browse()
measure <- 
structure(list(chest = c(34L, 37L, 38L, 36L, 38L, 43L,
40L, 38L, 40L, 41L, 36L, 36L, 34L, 33L, 36L, 37L, 34L, 36L, 38L,
35L), waist = c(30L, 32L, 30L, 33L, 29L, 32L, 33L, 30L, 30L, 32L,
24L, 25L, 24L, 22L, 26L, 26L, 25L, 26L, 28L, 23L), hips = c(32L,
37L, 36L, 39L, 33L, 38L, 42L, 40L, 37L, 39L, 35L, 37L, 37L, 34L,
38L, 37L, 38L, 37L, 40L, 35L), 
factor(rep(c("male", "female"), each = 10))), .Names = c("chest", "waist", "hips", "gender"), class = "data.frame", row.names = c(NA, -20L))


lab <- "ScatterPlotMatrix for measure data excluded Gender"
dat2 <- measure[, -4]
labs <- names(dat2)
par(mfrow = c(3, 3), oma = c(0, 0, 2, 0), mar = c(2, 5, 4, 2))
boxplot(x = dat2[, 1], col = "orange", main = labs[1], ylab = labs[1], 
        col.main = "darkblue", cex.main = 2, font.main = 2, 
        col.lab = "darkblue", cex.lab = 2, font.lab = 2, xlab = "")
bvbox(cbind(dat[, c(1, 2)]), pch = 16, col = "darkblue", 
      main = labs[2], col.main = "darkblue", cex.main = 2, font.lab = 2, 
      xlab = "", ylab = "")
bvbox(cbind(dat[, c(1, 3)]), pch = 16, col = "darkblue", main = labs[3],
      col.main = "darkblue", cex.main = 2, font.lab = 2, xlab = "", 
      ylab = "")
bvbox(cbind(dat[, c(2, 1)]), pch = 16, col = "darkblue", 
      ylab = labs[2], 
      col.lab = "darkblue", cex.lab = 2, font.lab = 2, xlab = "")
boxplot(x = dat[, 2], col = "orange")
bvbox(cbind(dat[, c(2, 3)]), pch = 16, col = "darkblue", xlab = "", 
      ylab = "")
bvbox(cbind(dat[, c(3, 1)]), pch = 16, col = "darkblue", ylab = labs[3], 
col.lab = "darkblue", cex.lab = 2, font.lab = 2, xlab = "")
bvbox(cbind(dat[, c(3, 2)]), pch = 16, col = "darkblue", xlab = "", 
      ylab = "")
boxplot(x = dat[, 3], col = "orange", xlab = "")
mtext(lab, outer = TRUE, cex = 1.5, font = 4, col=rgb(0.1,0.3,0.5,0.5))


####################################

attach(measure)
library(KernSmooth)
par(mfrow = c(3, 3), oma = c(0, 0, 2, 0), mar = c(2, 5, 4, 2))



hist(chest, axes = F, ylab = "")

res1 <-  bkde2D(measure[, 1:2], bandwidth = sapply(measure[, 1:2], dpik))
contour(res1$x1, res1$x2, res1$fhat)

res2 <-  bkde2D(measure[, c(1, 3)], bandwidth = sapply(measure[, c(1, 3)], dpik))
contour(res2$x1, res2$x2, res2$fhat)

res3 <-  bkde2D(measure[, 2:1], bandwidth = sapply(measure[, 2:1], dpik))
contour(res3$x1, res3$x2, res3$fhat)

hist(waist, axes = F, ylab = "")

res4 <-  bkde2D(measure[, c(2, 3)], bandwidth = sapply(measure[, c(2, 3)], dpik))
contour(res4$x1, res4$x2, res4$fhat)

res5 <-  bkde2D(measure[, c(3, 1)], bandwidth = sapply(measure[, c(3, 1)], dpik))
contour(res5$x1, res5$x2, res5$fhat)

res6 <-  bkde2D(measure[, c(3, 2)], bandwidth = sapply(measure[, c(3, 2)], dpik))
contour(res6$x1, res6$x2, res6$fhat)

hist(hips, axes = F, ylab = "")

detach(measure)
```








***
***


## Ex. 2.4 Construct a further scatterplot matrix of the body measurements data that labels each point in a panel with the gender of the individual, and plot on each scatterplot the separate estimated bivariate densities for men and women.


```{r fig.height = 9, fig.width = 9}
theme_update(plot.title = element_text(
color = "darkblue", size = 30, face = "bold"), 
axis.title.x = element_text(
color = "darkblue", size = 30, face = "bold"), 
axis.title.y = element_text(
color = "darkblue", size = 30, face = "bold"))
  
par(mfrow = c(3, 3), oma = c(0, 0, 2, 0), mar = c(2, 5, 4, 2))
old <- theme_set(theme_bw())

P1 <- measure %>%
ggplot(aes(x = chest, group = gender, colour = gender)) +
  geom_density() + labs(x = "", y = "chest", title = "chest") + 
  theme(legend.position = "bottom") 
measure %>%
mutate(Gen = case_when(gender == "female" ~ "f", 
TRUE ~ "m")) %>%
ggplot(aes(x = chest, y = waist, colour = gender)) + 
geom_point(shape = "") +
geom_text(aes(label = Gen)) -> P11
P2 <- measure %>%
  ggplot(aes(x = chest, y = waist, group = gender, colour = gender)) +
  geom_point(shape = "",  size = 5) + 
  geom_text(aes(label = gender)) + 
  geom_density_2d() + theme(legend.position = "bottom") + 
  labs(x = "", y = "", title = "waist")


P3 <- measure %>%
  ggplot(aes(x = chest, y = hips, group = gender, colour = gender)) +
  geom_point(shape = "",  size = 5) + 
  geom_text(aes(label = gender)) + 
  geom_density_2d() + theme(legend.position = "bottom") + 
  labs(x = "", y = "", title = "hips")


P4 <- measure %>%
  ggplot(aes(x = waist, y = chest, group = gender, colour = gender)) +
  geom_point(shape = "",  size = 5) + 
  geom_text(aes(label = gender)) + 
  geom_density_2d() + theme(legend.position = "bottom") + 
  labs(x = "", y = "waist", title = "")


P5 <- measure %>%
ggplot(aes(x = waist, group = gender, colour = gender)) +
  geom_density() + labs(x = "x", y = "y") + 
  theme(legend.position = "bottom") + 
  labs(x = "", y = "", title = "")


measure %>%
mutate(Gen = case_when(gender == "female" ~ "f", 
TRUE ~ "m")) %>%
ggplot(aes(x = chest, y = hips, colour = gender)) + 
geom_point(shape = "") +
geom_text(aes(label = Gen)) -> P12

P6 <- measure %>%
  ggplot(aes(x = waist, y = hips, group = gender, colour = gender)) +
  geom_point(shape = "",  size = 5) + 
  geom_text(aes(label = gender)) + 
  geom_density_2d() + theme(legend.position = "bottom") + 
  labs(x = "", y = "", title = "")


P7 <- measure %>%
  ggplot(aes(x = hips, y = chest, group = gender, colour = gender)) +
  geom_point(shape = "",  size = 5) + 
  geom_text(aes(label = gender)) + 
  geom_density_2d() + theme(legend.position = "bottom") + 
  labs(x = "", y = "hips", title = "")



P8 <- measure %>%
  ggplot(aes(x = hips, y = waist, group = gender, colour = gender)) +
  geom_point(shape = "",  size = 5) + 
  geom_text(aes(label = gender)) + 
  geom_density_2d() + theme(legend.position = "bottom") + 
  labs(x = "", y = "", title = "")


P9 <- measure %>%
ggplot(aes(x = hips, group = gender, colour = gender)) +
  geom_density() + labs(x = "", y = "") + 
  theme(legend.position = "bottom")

measure %>%
mutate(Gen = case_when(gender == "female" ~ "f", 
TRUE ~ "m")) %>%
ggplot(aes(x = waist, y = hips, colour = gender)) + 
geom_point(shape = "") +
geom_text(aes(label = "Gen")) -> P13

library(ggpubr)

Fig <- ggarrange(P11, P2, P3, P4, P12, P6, P7, P8, P13, 
                 nrow = 3, ncol = 3)


Fig <- annotate_figure(Fig, 
top = text_grob("ScatterPlot Matrix For Measure data", color = "plum4", 
face = "bold", size = 14), 

right = text_grob("this is bivariate density Plot and contour plot", 
color = "plum3", 
rot = 90, face = "bold", size = 14),
fig.lab = "made by Habib", fig.lab.face = "bold", 
fig.lab.pos = "bottom.left")
Fig
```







***
***





Ex. 2.5 Construct a scatterplot matrix of the chemical composition of
Romano-British pottery given in Chapter 1 (Table 1.3), identifying each
unit by its kiln number and showing the estimated bivariate density on
each panel. What does the resulting diagram tell you?


```{r fig.height = 10, fig.height = 10}
Name <- c("Al203", "Fe203", "Mg0", "Ca0", "Na20", 
          "K20", "Ti02", "Mn0", "Ba0", "Klin")
dat <- read.table(file = file.choose(), header = FALSE) 
names(dat) <- Name
dat <- dat %>%
  mutate(Klin = as.factor(Klin))
dim(dat)
head(dat)

dat$Klin %>% table





## second way

pairs(dat[, -10], 
      panel = function(x, y, ...){
      res <-  MASS :: kde2d(x, y)
      points(x, y,  ...)
      text(x, y, labels = dat$Klin)
     contour(res$x, res$y, res$z, add = TRUE)}
     , pch = ".", cex = 1.5)

```

***
***


## Ex. 2.6 Construct a bubble plot of the earthquake data using latitude and
longitude as the scatterplot and depth as the circles, with greater depths
giving smaller circles. In addition, divide the magnitudes into three equal
ranges and label the points in your bubble plot with a different symbol
depending on the magnitude group into which the point falls.





```{r fig.width = 7, fig.height = 7}

data(quakes)
names(quakes)

dat2 <- quakes

dat <- dat2 %>%
mutate(depth_fact = factor(cut_number(dat2$depth, 3))) 

dat$depth_fact %>% summary
dat %>%
  ggplot(aes(x = lat, y = long, size = depth, 
             shape = depth_fact, colour = depth_fact)) + 
  geom_point() + 
  labs(x = "lat", y = "long", title = "Bubble Plot for quakes data", 
       caption = "made by habib") + theme(legend.position = "bottom")
```








